\chapter{Power Markets and The Role of Flexibility Management}
\label{ch:market}
\textit{This chapter introduces some key concepts of power market elements and how the role of flexibility management is determined by them. We adopted a generalized method to extract the key attributes of power market structures that have impacts on value of flexibility management. The purpose of this chapter is to offer a qualitative analytic framework and}
%a comprehensive and comparative view on flexibility management in different power market regimes.}

\section{Power market frameworks: a comparative analysis}

Started in the 1980s and facilitated in 1990s, liberalized power markets has been the mainstream worldwide, especially in developed countries where the constructions of power infrastructure have been largely completed \cite{Srivastava2011,Ranci2013}.
Nowadays, there are many maturely existing liberalized power markets. However, since different preconditions exists in different countries due to historical, political and climatical reasons, the structure of their power markets tend to be very heterogeneous. Moreover, with the development of technologies, for instance the renewable penetration and rise of demand response as well as electricity storage, power markets face pending or undergoing restructuring, make them a rapidly changing field of the economy. \cite{Ziel2015}

These spatial and temporal variances bring great challenges to our study as the business models of flexiblity managment and values out of them depend extensively on the power market structure. Hereby we reviewed and analyzed the existing mechanisms of how power makets can possibly enable the value creation of flexibility management. Proposing novel market mechanisms is out of the scope of our study. 

%\subsection{General structure of power markets}

%\subsection{Key attributes of power market structure}


\section{Proposed business model}
Power exchange / Power pool

Capacity or not

Locational pricing or not
\subsection{Offering}

\subsection{Customer}

\subsection{Application}


Power exchange vs. power pool

The power exchange is a centralized market that usually uses simple offers/bids in the form of price-quantity pairs. The market operator (MO) solves a convex linear programming (LP) problem on an hour-by-hour basis to match the supply offers with the demand bids and determines the market clearing price without taking into account any technical aspects constraining the operation of the generating units or the transmission system. In a power exchange, the role of system operator (SO) is limited to preserving the system security, while each producer is responsible for self-scheduling his own units by solving a decentralized price-based unit commitment (PBUC).

On the other hand, a power pool is a centrally organized market that usually uses complex unit offers and the independent system operator (ISO, who usually acts both as MO and SO) solves a non-convex optimization problem, where a simultaneous 24-hour co-optimization of energy and reserve resources is performed under a large set of unit and system constraints (e.g., unit start-up and shut-down procedures, minimum-up/down time constraints, min/max power output restrictions, ramp-rate limits, transmission limits etc.). In a power pool, the objective of the ISO is the minimization of the system total production cost through a centralized unit commitment. The producers must follow the commitment schedule and the dispatch instructions issued by the ISO in exchange of the make-whole side payments that guarantee that they will fully recover their total operating costs.

%\section{Power market design and structure}
%\subsection{PJM}

%\subsection{Germany}

%\subsection{Australia}
%%\subsection{Day-ahead energy market}

%\subsection{Real-time balancing energy market}

%\subsection{Regulation market}

%\subsection{Synchronized, non-synchronized and supplementary reserves market}

%\subsection{Overview of other non-market services}

%\section{Power market in Germany}
%\subsection{Day-ahead energy market}

%\subsection{Intraday balancing energy market}

%\subsection{Primary, secondary and tertiary frequency control markets}

%\subsection{Overview of other non-market services}

%\section{Power market in  Australia}
%\subsection{Energy market}

%\subsection{Frequency control ancillary services market}

%\subsection{Overview of other non-market services}

%\input{PJM}

%\section{Regulatory and market framework for flexibility resourses}